Think deep for reef resilience

### 2025-09-12 paper/paper_draft.qmd
Changed text to be more direct to shorten

To apply exploratory scenario modelling, we employed a framework involving two separate
systems: ADRIA (the Adaptive Dynamic Reef Intervention Algorithms platform) and CoralBlox.
ADRIA is a multi-criteria decision support platform designed to help reef managers address
climate change challenges in coral reef systems [@iwanaga2023]. The ADRIA platform provides
functionality for scenario generation and analysis as well as providing the technical
infrastructure to couple environmental and ecosystem representations. CoralBlox is a
computationally efficient coral population model. It simulates functional/morphological
groups of corals over their life cycles, competing for available space over time
[@ribeirodealmeida2024]. The coupled framework is referred to as ADRIA-CoralBlox.

The coral demographic representation and associated model implementation are further
detailed in the supplementary material. Of important note, however, is that the
representation of bleaching mortalities followed Bozec et al., [-@bozec2022] albeit with a
scaling coefficient to incorporate the amelioration effect that depth may have, following
Baird et al., [@baird2018]. This depth-scaling reduces the proportion of coral colonies
impacted by high heat stress as depth increases. The impacts of tropical cyclones, wave
activity, Crown of Thorns Starfish or coral disease were not included in this investigation,
in order to focus on assessing reef resilience under possible future warming scenarios.

ADRIA-CoralBlox was run for 4096 scenarios for each GCM (20,480 total model runs). Here, a
scenario represents a unique combination of assumed environmental (namely one of the five
DHW trajectories) and ecological factors including coral growth and mortality rates, and
coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol' sampling
method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi Monte Carlo
approach which balances scenario exploration with improved convergence properties compared
to pseudo-random sampling. A detailed comparative analysis of the sampling methods can be
found in [@kucherenko2015]. For the purpose of scenario assessment, the resulting coral
cover time series for each model run were scaled by the total area of each reef to represent
cover as a percentage of the total estimated reef area. The scenario-median timeseries for
each reef was used for further analyses. These time series of coral cover were then
clustered based on their temporal behaviour.

>>>>

Two separate systems were employed to explore scenarios: ADRIA (the Adaptive Dynamic Reef
Intervention Algorithms platform) and CoralBlox. ADRIA is a multi-criteria decision support
platform designed for the assessment and analysis of coral reef intervention strategies
under uncertainty [@iwanaga2023]. The ADRIA platform provides functionality for scenario
generation and analysis as well as providing the technical infrastructure to couple
environmental and ecosystem representations. CoralBlox is a computationally efficient coral
population model. It simulates functional/morphological groups of corals over their life
cycles, competing for available space over time [@ribeirodealmeida2024]. The coupled
framework is referred to as ADRIA-CoralBlox.

The coral demographic representation and associated model implementation are further
detailed in the supplementary material. Of important note is that the representation of
bleaching mortalities followed Bozec et al., [-@bozec2022] albeit with a scaling coefficient
to incorporate the amelioration effect that depth may have, following Baird et al.,
[@baird2018]. This depth-scaling reduces the proportion of coral colonies impacted by high
heat stress as depth increases. The impacts of tropical cyclones, wave activity, Crown of
Thorns Starfish or coral disease were not included in this investigation, in order to focus
on assessing reef resilience under possible future warming scenarios.

ADRIA-CoralBlox was run for 4096 scenarios for each GCM (20,480 total model runs). Here, a
scenario represents a unique combination of assumed environmental (namely one of the five
DHW trajectories) and ecological factors including coral growth and mortality rates, and
coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol' sampling
method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a randomised Quasi Monte
Carlo approach which balances scenario exploration with improved convergence properties
compared to pseudo-random sampling. A detailed comparative analysis of sampling methods can
be found in [@kucherenko2015]. For the purpose of scenario assessment, the resulting coral
cover time series for each model run were scaled by the total area of each reef to represent
cover as a percentage of the total estimated reef area. The scenario-median timeseries for
each reef was used for further analyses. These time series of coral cover were then
clustered based on their temporal behaviour.
